---
title: "APTP Clean"
author: "Cat"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(lubridate)
library(readxl)
library(sf)
library(estimatr)
library(ggthemes)
library(hrbrthemes)
library(plotly)

setwd("~/Documents/Fall Data Science/R Code/APTP-Project")


opd_data_raw <- read_csv("OPD_PublicCallData_2019.csv")
opd_grouped <- read_excel("Oakland Call Data Grouping by Description.xlsx")
shape1.sf <- st_read("geo_export_003c8e85-82f8-4ef8-aec4-227e86a03525.shp")

```

## R Markdown

```{r response times}

opd_data_raw  %>% 
  mutate(response_time = `Arrival Time` - `Transmit Time`) -> opd_data_new



```

There is an issue here where this join is creating duplicates, I would love to figure this out in the merge but for now I'm just going to dedup separately

```{r merging on grouping}

opd_data_new %>% 
  left_join(opd_grouped, "Description", na_matches = "never") %>% 
  select(Code:response_time, Grouping, 'Violent Crime' ) -> opd_data_new


distinct(opd_data_new) -> opd_data_new
```



```{r dispatched calls}

opd_data_new %>% 
   mutate(dispatched = !is.na(`Dispatch Time`)) -> opd_data_new
```

```{r num calls by source and by grouping}
opd_data_new %>% 
  group_by(Source) %>% 
  summarize(count = n()) %>% 
  mutate(perc = round(count/sum(count), digits = 4)) %>% 
  arrange(desc(count))


opd_data_new %>% 
  group_by(Grouping) %>% 
  summarize(count = n()) %>% 
  mutate(perc = round(count/sum(count), digits = 4)) %>% 
  arrange(desc(count))

```

To work on later - there are some calls here that are missing Grouping that I excluded - I should go back and get them later. I also want to better investigate whether my assumptions about whether police (versus someone else) were dispatched are correct.

```{r overall police activity perc violent crime}

opd_data_new %>% 
  filter(!is.na(Grouping), dispatched = TRUE) %>% 
  group_by(`Violent Crime`) %>% 
  summarize(count = n()) %>% 
  mutate(perc = round(count/sum(count), digits = 4)) %>% 
  arrange(desc(count)) -> pie_chart_perc_violent_crime


police_activity <- plot_ly(pie_chart_perc_violent_crime, labels = ~`Violent Crime`, values = ~perc, type = 'pie')
police_activity <- police_activity %>% layout(title = 'Police Activity Related to Violent or Serious Crime',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

police_activity
```


```{r call types by source}

opd_data_new %>% 
  filter(Source == '911 CALL') %>% 
  group_by(Grouping, `Violent Crime`) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(perc = round(count/sum(count), digits = 4)) %>% 
  arrange(desc(count))


opd_data_new %>% 
  filter(Source == 'DEFAULT - PHONE INIT') %>% 
  group_by(Grouping, `Violent Crime`) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(perc = round(count/sum(count), digits = 4)) %>% 
  arrange(desc(count))

opd_data_new %>% 
  filter(Source %in% c('FIELD INITIATED BY D', 'MDT-INITIATED')) %>% 
  group_by(Grouping, `Violent Crime`) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(perc = round(count/sum(count), digits = 4)) %>% 
  arrange(desc(count))

```

```{r visualize 911 calls}

opd_data_new %>%
  filter(Source == "911 CALL") %>% 
  filter(!is.na(Grouping), !is.na(`Violent Crime`)) %>% 
  group_by(Grouping, `Violent Crime`) %>% 
  summarize(count = n())  %>% 
  ungroup() %>% 
  mutate(percent = count/sum(count)*100) -> nine91_calls


nine91_calls %>%
  filter(!is.na(Grouping), !is.na(`Violent Crime`), percent >= .2) %>% 
  ggplot(aes(x = percent, y = reorder(Grouping, percent), fill = `Violent Crime`)) +
    geom_bar(stat = "identity") +
    theme_ipsum()+
    labs(title = "Oakland 2019 911 Calls by Type", caption = '2019 OPD Call Data, N = 442,839') +
    theme(axis.title.y=element_blank()) +
    scale_fill_brewer(palette="Dark2")

```

```{r violent crime pie chart}


fig <- plot_ly(nine91_calls, labels = ~`Violent Crime`, values = ~percent, type = 'pie')
fig <- fig %>% layout(title = '911 Calls Related to Violent or Serious Crime',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig
```


```{r visualize all calls}

opd_data_new %>%
  group_by(Grouping, `Violent Crime`) %>% 
  summarize(count = n())  %>% 
  filter(count >= 200, !is.na(Grouping)) %>% 
  ggplot(aes(x = count, y = reorder(Grouping, count), fill = `Violent Crime`)) +
    geom_bar(stat = "identity") +
    theme_ipsum()



```
'


```{r average response times for 911 calls}

opd_data_new %>% 
   mutate(response_time_hms = hms::as_hms(response_time)) -> opd_data_new

opd_data_new %>% 
  mutate(response_time_minutes = hour(response_time_hms)*60 + minute(response_time_hms) + 1) -> opd_data_new

opd_data_new %>% 
  filter(Source == "911 CALL", !is.na(response_time_minutes), !is.na(Grouping)) %>% 
  group_by(Grouping) %>% 
  summarize(mean_response = mean(response_time_minutes),
            count = n()) %>% 
  arrange(mean_response) -> response_time_table

write.csv(response_time_table,"response_time_table.csv", row.names = FALSE)




response_time_table %>%
  ggplot(aes(x = mean_response, y = reorder(Grouping, -mean_response), fill = desc(`mean_response`))) +
    geom_bar(stat = "identity") +
    scale_colour_gradient(low = "tomato1", high = "red4") +
    theme_ipsum()+
    labs(title = "Response Time by Call Type", caption = '2019 OPD Call Data, N = 442,839') +
    theme(axis.title.y=element_blank())  +
    theme(axis.title.x=element_blank())  + 
    theme(legend.position = "none")

```


```{r average response times for 911 calls by beat}
library(RColorBrewer)

opd_data_new %>% 
  filter(Source == "911 CALL", !is.na(response_time_minutes)) %>% 
  group_by(Beat) %>% 
  summarize(mean_response = mean(response_time_minutes),
            count = n()) %>% 
  arrange(mean_response) -> mean_response_beat


opd_data_new %>% 
  filter(Source == "911 CALL", !is.na(response_time_minutes)) %>% 
  group_by(Beat) %>% 
  summarize(mean_response = mean(response_time_minutes),
            count = n()) -> scatter_plot

opd_data_new %>% 
  filter(Source == "911 CALL", !is.na(response_time_minutes)) %>% 
  group_by(Beat) %>% 
  summarize(mean_response = mean(response_time_minutes),
            count = n()) %>% 
  ggplot(aes(x = Beat, y = mean_response, size = count, color = mean_response)) +
  geom_point()  +
    scale_colour_gradient(low = "turquoise1", high = "blue4") +
    theme_ipsum() +
    labs(color = "Time in Minutes", size = 'Number of Calls', y = 'Time in Minutes') +
    theme(axis.text.x = element_text(angle = 90))+
    labs(title = "Response Time by Beat", caption = '2019 OPD Call Data') 
    theme(plot.caption=element_text(face="italic")) 

```


```{r average response times for 911 calls by beat in plotly}



fig <- plot_ly(data = scatter_plot, type = "scatter", x = ~Beat, y = ~mean_response, size = ~count, color = ~mean_response)


fig
```


```{r average response times for 911 calls by beat for violent crimes}


opd_data_new %>% 
  filter(Source == "911 CALL", !is.na(response_time_minutes), `Violent Crime` = TRUE) %>% 
  group_by(Beat) %>% 
  summarize(mean_response = mean(response_time_minutes),
            count = n()) %>% 
  arrange(mean_response)




opd_data_new %>% 
  filter(Source == "911 CALL", !is.na(response_time_minutes), `Violent Crime` = TRUE) %>% 
  group_by(Beat) %>% 
  summarize(mean_response = mean(response_time_minutes),
            count = n()) %>% 
  ggplot(aes(x = Beat, y = mean_response, size = count, color = desc(mean_response))) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90))  +
  theme_ipsum() 

```

```{r cleaning data for merge}

shape1.sf %>% 
  left_join(mean_response_beat, c("name" = "Beat"), na_matches = "never") %>% 
  select_all() -> shape_response.sf


```


```{r mapping response times by beat}


theme_map <- function(base_size=9, base_family="") {
    require(grid)
    theme_bw(base_size=base_size, base_family=base_family) %+replace%
        theme(axis.line=element_blank(),
              axis.text=element_blank(),
              axis.ticks=element_blank(),
              axis.title=element_blank(),
              panel.background=element_blank(),
              panel.border=element_blank(),
              panel.grid=element_blank(),
              panel.spacing=unit(0, "lines"),
              plot.background=element_blank(),
              legend.justification = c(0,0),
              legend.position = c(0,0)
              )
}

shape_response.sf %>% 
  ggplot(aes(fill = mean_response)) + 
  geom_sf() + 
  coord_sf() +
  theme_map()   +
    labs(title = "Average 911 Response Time by Beat", caption = '2019 OPD Call Data') +
    theme(plot.caption=element_text(face="italic")) +
    labs(fill = "Time in Mins")+
    scale_fill_gradient(low = "turquoise1", high = "blue4") 

```

```{r regression}


lmout <- lm_robust(response_time ~ Beat + factor(Priority), data = opd_data_new)
summary(lmout)

lmout2 <- lm_robust(response_time ~ Beat + Grouping + factor(Priority), data = opd_data_new)
summary(lmout2)

lmout3 <- lm_robust(response_time ~ Beat + Grouping, data = opd_data_new)
summary(lmout3)


```

